News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-05 rupnik2015.tex
updates

As input, we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori, we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events. The task is divided into two parts: detecting events within each language and then linking events across languages. In this paper we address the second step.

As input we consider a high volume of articles in different languages. By using a language detector the stream is split into separate monolingual streams. Within each monolingual stream, an online clustering approach is employed, where tracked clusters correspond to our definition of events - this is based on the Event Registry system \shortcite{Leban2014W,Leban2014I}. Our main goal in this paper is to connect such clusters (representations of events) across languages, that is detect that a set of articles in language $A$ reports on the same event as a set of articles in language $B$.

@@

Since articles in a cluster are in a single language, we also want to identify any other existing clusters that report about the same event in other languages and join these clusters into the same event. This task is performed using a classification approach that is the second major contribution of this paper. It is described in detail in Section~\ref{sec:linking}.

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both step rely heavily on simliarity computation, which implies that this must be highly scalable.

Therefore, we focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This  simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpora (as opposed to parallel translation-based corpora), with missing data. Supporting a new language should be straightforward.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day. Computing the similarity between a new document and a set of known documents should be efficient (the main application is linking documents between two monolingual streams).
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

@@

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. It can be interpreted as using the sample covariance matrix between features of two languages to define the dot product that is used to compute similarities.
The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as opposed to 10,000 document limit reported in \cite{ESA}). In addition, ESA is less suited for computing similarities between two large monolingual streams. For example, each day we have to compute similarities between 500,000 English and 500,000 German news articles. Comparing each German news article with 500,000 English news articles is either prohibitively slow (involves projecting all English articles on Wikipedia) or consumes too much memory (involves storing the projected English articles, which for a Wikipedia of size 1,000,000 is a 500,000 by 1000,0000 non-sparse matrix).

@@

A system that is significantly different, but worth mentioning is the GDELT project~\cite{Leetaru2013Gdelt}. In GDELT, events are also extracted from articles, but in their case, an event is specified in a form of a triple containing two actors and a relation. The project contains an extensive vocabulary of possible relations, mostly related to political events. In order to identify events, GDELT collects articles in more than 65 languages and uses machine translation to translate them to English. All information extraction is then done on the translated article.

@@

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare douments in a mono-lingual setting. We then define a way to measure cross-lingual similarity that is natural for the models we consider.

@@

The cross-lingual similarity models presented in this paper are based on comparable corpora. That is, a corpus of documents in multiple languages, with alignment between documents that are of the same topic, or even a rough translation of each other. Wikipedia is an example of a comparable corpus, where a specific entry can be described in multiple languages (e.g. ``Berlin" is currently described in 222 languages). News articles represent another example, where the same event can be described by newspapers in several languages.

@@

Each article $a \in A$ is written in a language $\ell$, where $\ell \in L = \{\ell_1,\ell_2,...,\ell_m\}$. For each language $\ell$, we obtain a set of monolingual clusters $C_{\ell}$. That is, all articles from each cluster $c \in C_{\ell}$ are in language $\ell$. Given a pair of languages $\ell_a \in L$, $\ell_b \in L$ and $\ell_a \not= \ell_b$, we would like to identify all cluster pairs $(c_i, c_j) \in C_{\ell_a} \times C_{\ell_b}$ such that $c_i$ and $c_j$ describe the same event.

@@

{
\renewcommand\tabcolsep{3pt}
\begin{table}[h!]
\centering
\caption{Pairwise training -- test sizes (in thousands)
The first row represents the size of the training sets used to construct the mappings in low-dimensional language independent space using the English--\emph{en} as a hub. The diagonal elements represent the number of the unique training documents and test documents in each language.
}
\label{table:train_test}
{
\small
\begin{tabular}{c|c|c|c|c|c|c|c|c|}
&	en&	es&	ru&	sl&	hi&	war&	ht&	pms\\\cline{1-9}
en&	671~-~4.64&	463~-~4.29&	369~-~3.19&	50.3~-~2&	14.4~-~2.76&	8.58~-~2.41&	 17~-~2.32&	16.6~-~2.67\\
\cline{2-9}
es&	\multicolumn{1}{c|}{}	&	463~-~4.29&	187~-~2.94&	28.2~-~1.96&	8.72~-~2.48&	 6.88~-~2.4&	13.2~-~2&	 13.8~-~2.58\\
\cline{3-9}
ru&	\multicolumn{2}{c|}{}	&	369~-~3.19&	29.6~-~1.92&	9.16~-~2.68&	2.92~-~1.1&	 3.23~-~2.2&	10.2~-~1.29\\
\cline{4-9}
sl&	\multicolumn{3}{c|}{}	&	50.3~-~2&	3.83~-~1.65&	1.23~-~0.986&	0.949~-~1.23&	 1.85~-~0.988\\
\cline{5-9}
hi&	\multicolumn{4}{c|}{}	&	14.4~-~2.76&	0.579~-~0.76&	0.0~-~2.08&	0.0~-~0.796\\
\cline{6-9}
war&	\multicolumn{5}{c|}{}	&	8.58~-~2.41&	0.043~-~0.534&	0.0~-~1.97\\
\cline{7-9}
ht&	\multicolumn{6}{c|}{}	&	17~-~2.32&	0.0~-~0.355\\
\cline{8-9}
pms&	\multicolumn{7}{c|}{}	&	16.6~-~2.67\\
\cline{9-9}
\end{tabular}
}
\end{table}
}

We further inspect the properties of the training sets by roughly estimating the fraction \texttt{rank(A)/min(size(A))} for each English training matrix and its corresponding mate matrix. Ideally, these two fractions are approximately the same so  both aligned spaces should have reasonably similar dimensionality. We display these numbers as pairs in Table \ref{table:rank}.

\begin{table}[h]
\caption{Dimensionality drift}
\label{table:rank}
\begin{tabular}{|c|c|c|c|c|c|c|}
\cline{1-7}
(en, de)     &   (en, ru)     &   (en, sl)       &     (en, hi)&   (en, war)      &      (en, ht) &   (en, pms)\\
\cline{1-7}
(0.81, 0.89)   &  (0.8, 0.89)  &   (0.98, 0.96)    &    (1, 1)  &   (0.74, 0.56)  &      (1, 0.22)  &   (0.89, 0.38)\\
\cline{1-7}
\end{tabular}
\end{table}

@@

In Section~\ref{sec:models}, we  described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and hub CCA. As a training set, we used common Wikipedia alignment for all three languages. To test which of these algorithms performed best, we made the following test. For each of the three algorithms, we analyzed all articles in Event Registry and for each article computed the most similar articles in other languages. To test how informative the identified similar articles are for cluster linking we then trained three classifiers as described in Section~\ref{algo:features} -- one for each algorithm. Each classifier was allowed to use as learning features only the cross-lingual article linking features for which values are determined based on the selected algorithm ($k$-means, LSI and hub CCA). The results of the trained models are shown in Table~\ref{table:linkingEvalAlgos}.

@@

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15,000$ aligned test documents and using $300,000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. The hub CCA approach with $500$ topic vectors manages to perform comparably to the $LSI$-based approach with $1,000$ topic vectors, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

@@

\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1,000 topic vectors obtained from different cross-lingual similarity algorithms. The table shows for each of the algorithms the obtained classification accuracy, precision and recall.}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & A & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

>>>>

As input, we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori, we do not know the coverage of the articles, that is, not all the events may be covered and we do not know that all the articles necessarily fit into one of the events. The task is divided into two parts: detecting events within each language and then linking events across languages. In this paper we address the second step.

As input we consider a high volume of articles in different languages. By using a language detector the stream is split into separate monolingual streams. Within each monolingual stream, an online clustering approach is employed, where tracked clusters correspond to our definition of events - this is based on the Event Registry system \shortcite{Leban2014W,Leban2014I}. Our main goal in this paper is to connect such clusters (representations of events) across languages, that is, to detect that a set of articles in language $A$ reports on the same event as a set of articles in language $B$.

@@

Since articles in a cluster are in a single language, we also want to identify any other existing clusters that report about the same event in other languages and join these clusters into the same event. This task is performed using a classification approach which is the second major contribution of this paper. It is described in detail in Section~\ref{sec:linking}.

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both step rely heavily on similarity computation, which implies that this must be highly scalable.

Therefore, we focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpora (as opposed to parallel translation-based corpora), with missing data.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day. Computing the similarity between a new document and a set of known documents should be efficient (the main application is linking documents between two monolingual streams).
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

@@

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices whose dimensions are a product of the training set size and the vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Cross-Lingual Explicit Semantic Analysis (CL-ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. It can be interpreted as using the sample covariance matrix between features of two languages to define the dot product which is used to compute similarities.
The authors of CL-ESA compare it to CL-LSI and find that CL-LSI can outperform CL-ESA in an information retrieval, but is costlier to optimize over a large corpus (CL-ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as opposed to 10,000 document limit reported in \cite{ESA}). In addition, CL-ESA is less suited for computing similarities between two large monolingual streams. For example, each day we have to compute similarities between 500,000 English and 500,000 German news articles. Comparing each German news article with 500,000 English news articles is either prohibitively slow (involves projecting all English articles on Wikipedia) or consumes too much memory (involves storing the projected English articles, which for a Wikipedia of size 1,000,000 is a 500,000 by 1000,0000 non-sparse matrix).

@@

A system, which is significantly different but worth mentioning, is the GDELT project~\cite{Leetaru2013Gdelt}. In GDELT, events are also extracted from articles, but in their case, an event is specified in a form of a triple containing two actors and a relation. The project contains an extensive vocabulary of possible relations, mostly related to political events. In order to identify events, GDELT collects articles in more than 65 languages and uses machine translation to translate them to English. All information extraction is then done on the translated article.

@@

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare douments in a mono-lingual setting. We then define a way to measure cross-lingual similarity which is natural for the models we consider.

@@

The cross-lingual similarity models presented in this paper are based on comparable corpora. A \emph{comparable corpus} is a collection of documents in multiple languages, with alignment between documents that are of the same topic, or even a rough translation of each other. Wikipedia is an example of a comparable corpus, where a specific entry can be described in multiple languages (e.g. ``Berlin" is currently described in 222 languages). News articles represent another example, where the same event can be described by newspapers in several languages.

@@

Each article $a \in A$ is written in a language $\ell$, where $\ell \in L = \{\ell_1,\ell_2,...,\ell_m\}$. For each language $\ell$, we obtain a set of monolingual clusters $C_{\ell}$. More precisely, the articles corresponding to each cluster $c \in C_{\ell}$ are written in the language $\ell$. Given a pair of languages $\ell_a \in L$, $\ell_b \in L$ and $\ell_a \not= \ell_b$, we would like to identify all cluster pairs $(c_i, c_j) \in C_{\ell_a} \times C_{\ell_b}$ such that $c_i$ and $c_j$ describe the same event.

@@

{
\renewcommand\tabcolsep{3pt}
\begin{table}[h!]
\centering
\caption{Training -- test sizes (in thousands).
The first row represents the size of the training sets used to construct the mappings in low-dimensional language independent space using the English--\emph{en} as a hub. The diagonal elements represent the number of the unique training documents and test documents in each language.
}
\label{table:train_test}
{
\small
\begin{tabular}{c|c|c|c|c|c|c|c|c|}
&	en&	es&	ru&	sl&	hi&	war&	ht&	pms\\\cline{1-9}
en&	671~-~4.64&	463~-~4.29&	369~-~3.19&	50.3~-~2&	14.4~-~2.76&	8.58~-~2.41&	 17~-~2.32&	16.6~-~2.67\\
\cline{2-9}
es&	\multicolumn{1}{c|}{}	&	463~-~4.29&	187~-~2.94&	28.2~-~1.96&	8.72~-~2.48&	 6.88~-~2.4&	13.2~-~2&	 13.8~-~2.58\\
\cline{3-9}
ru&	\multicolumn{2}{c|}{}	&	369~-~3.19&	29.6~-~1.92&	9.16~-~2.68&	2.92~-~1.1&	 3.23~-~2.2&	10.2~-~1.29\\
\cline{4-9}
sl&	\multicolumn{3}{c|}{}	&	50.3~-~2&	3.83~-~1.65&	1.23~-~0.986&	0.949~-~1.23&	 1.85~-~0.988\\
\cline{5-9}
hi&	\multicolumn{4}{c|}{}	&	14.4~-~2.76&	0.579~-~0.76&	0.0~-~2.08&	0.0~-~0.796\\
\cline{6-9}
war&	\multicolumn{5}{c|}{}	&	8.58~-~2.41&	0.043~-~0.534&	0.0~-~1.97\\
\cline{7-9}
ht&	\multicolumn{6}{c|}{}	&	17~-~2.32&	0.0~-~0.355\\
\cline{8-9}
pms&	\multicolumn{7}{c|}{}	&	16.6~-~2.67\\
\cline{9-9}
\end{tabular}
}
\end{table}
}

We further inspect the properties of the training sets by roughly estimating the fraction \texttt{rank(A)/min(rows(A), cols(A))} for each English training matrix and its corresponding mate matrix, where $rows(A)$ and $cols(A)$ denote the number of rows and columns respectively. The denominator represents the theoretically highest possible rank the matrix $A$ could have. Ideally, these two fractions should be approximately the same - both aligned spaces should have reasonably similar dimensionality. We display these numbers as pairs in Table \ref{table:rank}.

\begin{table}[h]
\caption{Dimensionality drift. Each column corresponds to a pair of aligned corpus matrices between English and another language. The numbers represent the ratio between the numerical rank and the highest possible rank. For example, the column $en -- ht$ tells us that for the English-Creole pairwise-aligned corpus matrix pair, the English counterpart has full rank, but the Creole counterpart is far having full rank.}
\label{table:rank}
\begin{tabular}{|c|c|c|c|c|c|c|}
\cline{1-7}
en -- de     &   en -- ru     &   en -- sl       &     en -- hi &   en -- war      &      en -- ht &   en -- pms\\
\cline{1-7}
0.81 -- 0.89   &  0.8 -- 0.89  &   0.98 -- 0.96    &    1 -- 1  &  0.74 -- 0.56  &      1 -- 0.22  &   0.89 -- 0.38\\
\cline{1-7}
\end{tabular}
\end{table}

@@

In Section~\ref{sec:models}, we  described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and hub CCA. As a training set, we used common Wikipedia alignment for all three languages. To test which of these algorithms performed best, we made the following test. For each of the three algorithms, we analyzed all articles in Event Registry and for each article computed the most similar articles in other languages. To test how informative the identified similar articles are for cluster linking we then trained three classifiers as described in Section~\ref{algo:features} -- one for each algorithm. Each classifier was allowed to use as learning features \textbf{only} the cross-lingual article linking features for which values are determined based on the selected algorithm ($k$-means, LSI and hub CCA). The results of the trained models are shown in Table~\ref{table:linkingEvalAlgos}. We also show how the number of topics (the dimensions of the latent space) influences the quality, except in the case of the $k$-means algorithm, where only the performance on 500 topic vectors is reported, due to higher computational cost.

@@

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15,000$ aligned test documents and using $300,000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and $k$-means lags far behind when testing on Wikipedia data. The hub CCA approach with $500$ topic vectors manages to perform comparably to the $LSI$-based approach with $1,000$ topic vectors, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

@@

\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1,000 topic vectors obtained from different cross-lingual similarity algorithms. The table shows for each of the algorithms the obtained classification accuracy, precision and recall.}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & Accuracy & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}
